Quantifying maternal reproductive output of chondrichthyan fishes

### 2024-08-25 README.qmd
stuff

data <- read_csv(here::here("data", "empirical-plumbeus.csv"))

@@

## Run model

Compile 3 parameter logistic model

@@

>>>>

data <- read_csv(here::here("data", "empirical-plumbeus.csv")) %>% 
  select(-source)

@@

## 3 Parameter logistic model

Compile the 3 parameter logistic model. The model

@@

```{r}

# Parameter list
pars <- list(m50 = 160, m95 = 170, c = 0.5)

# Create objective function
obj <- MakeADFun(
  data = data,
  parameters = pars,
  DLL = "logistic3"
)

# Optimize
opt <- nlminb(obj$par, obj$fn, obj$gr, lower = c(0, 0, 0), upper = c(Inf, Inf, 1))
```

### 2024-08-25 rmds/manuscript.tex
stuff

\[
\hat{C}_{l} = \sum_a(F_{l}/Z_{l})(1-\exp(-Z_{l}))N_{l,a}
\]

@@

When using the 3PLF methods model performance typically declined as the periodicity of the reproductive cycle increased. This is seen in the positive correlation between \(\hat{L'_{50}}\) and \(\hat{P_{Max}}\), whereby true underlying \(P_{Max}\) affected the magnitude and direction of bias in these parameters (Figure 2). With lower values of \(P_{Max}\) it appeared to be more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote. As true underlying \(P_{Max}\) decreased there was also a greater potential for positive bias in \(\hat{P_{Max}}\). Additionally, longer reproductive cycles resulted in fewer maternal females, which was associated with larger bias (Figure 2).

@@

While the 3PLF-fixed method was in some cases able to reduce bias at lower sample sizes, a trade off in the use of this method was poorer interval coverage (Figure 4, Figure S6). For most simulations, interval coverage for \(\hat{L'_{50}}\) was well below the expected level of 50\%, irrespective of sample size. In contrast, for the 3PLF-estimated method interval coverage oscillated around 50\% for both \(\hat{L'_{50}}\) and \(\hat{P_{Max}}\).

@@

Maternal data were re-analysed for 1087 sandbar sharks including 640 mature individuals of which 32\% were in maternal condition. Using the 3PLF-estimated method, the maximum likelihood estimate for \(\hat{P_{Max}}\) was 0.48 (Table 2). Despite having a sample size of \textgreater{} 600 mature females, the proportion of maternal individuals at length was still uncertain and \(\hat{P_{Max}}\) was estimated to lie between 0.39 and 0.60 with 95\% confidence (Figure 7). For 3PLF-fixed models a value of \(P_{Max} = 0.5\) was strongly supported (\(\Delta_{AIC}\) = 14.15) over the alternative fixed value of \(P_{Max} = 0.33\). The model with \(P_{Max} = 0.5\) also outperformed the 3PLF-estimated method (\(\Delta_{AIC}\) = 1.86). Given the fixed model had one fewer estimated parameter, both models had essentially the same level of support given the data (Burnham and Anderson 2002).

@@

The empirical analysis carried out using data for the sandbar shark illustrates how novel insights can be gained from maternal data by estimating \(P_{Max}\). Detailed reproductive studies of the western North Atlantic population by Baremore and Hale (2012) and Piercy et al. (2016) were ambiguous about the frequency of reproduction. In both studies, the low proportion of pregnant females and bimodality in ovarian follicle size led the authors to conclude that the reproductive cycle was most commonly triennial or possibly longer at the population level. Reanalysis of these combined data sets using the 3PLF-estimated method, however, resulted in a maximum likelihood estimate of \(\hat{P_{Max}} = 0.48\), considerably larger than 0.37 - the proportion of pregnant mature females observed by Baremore and Hale (2012). Similarly, based on AIC values, a fixed value of \(P_{Max} = 0.5\) had a much higher relative plausibility than a value of 0.33 (Table 2).

This unexpected result may be related to the gradual attainment of asymptotic maternal status. Maturity in sandbar sharks appears to occur over an extended size range. \(L_{5}\) and \(L_{95}\) are 146cm and 176cm -- a length interval that corresponds to around seven years given the sandbar shark's slow growth rate (Hale and Baremore 2013). With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently. The outcomes of this study suggest that larger sharks may ultimately be capable of reproducing biennially. Given that few females appear to attain this size (Hale and Baremore 2013). this suggests the modal reproductive frequency likely still at least triennial. Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management; not only are larger females more fecund (Baremore and Hale 2012), they also reproduce more frequently. This implies that size selective fishing or the removal of larger and older female sharks could have a disproportionately greater impact on population productivity.

@@

@@

\bibitem[\citeproctext]{ref-higgs_reproductive_2020}
Higgs, J.M., Hoffmayer, E.R., Sulikowski, J.A., Driggers, W.B., Stiller, D.A., and Hendon, J.M. 2020. Reproductive biology of the finetooth shark (\emph{{Carcharhinus} isodon}) in the northern {Gulf} of {Mexico}, with evidence of both annual and biennial reproduction. Mar. Freshwater Res. Available from \url{https://doi.org/10.1071/MF20120}.

@@

\bibitem[\citeproctext]{ref-huveneers_future_2018}
Huveneers, C., Apps, K., Becerril-García, E.E., Bruce, B., Butcher, P.A., Carlisle, A.B., Chapple, T.K., Christiansen, H.M., Cliff, G., Curtis, T.H., Daly-Engel, T.S., Dewar, H., Dicken, M.L., Domeier, M.L., Duffy, C.A.J., Ford, R., Francis, M.P., French, G.C.A., Galván-Magaña, F., García-Rodríguez, E., Gennari, E., Graham, B., Hayden, B., Hoyos-Padilla, E.M., Hussey, N.E., Jewell, O.J.D., Jorgensen, S.J., Kock, A.A., Lowe, C.G., Lyons, K., Meyer, L., Oelofse, G., Oñate-González, E.C., Oosthuizen, H., O'Sullivan, J.B., Ramm, K., Skomal, G., Sloan, S., Smale, M.J., Sosa-Nishizaki, O., Sperone, E., Tamburin, E., Towner, A.V., Wcisel, M.A., Weng, K.C., and Werry, J.M. 2018. Future research directions on the {``{Elusive}''} white shark. Frontiers in Marine Science \textbf{5}: 455. doi: \href{https://doi.org/10.3389/fmars.2018.00455}{10.3389/fmars.2018.00455}.

@@

\bibitem[\citeproctext]{ref-montealegre-quijano_sexual_2014}
Montealegre-Quijano, S., Cardoso, A.T.C., Silva, R.Z., Kinas, P.G., and Vooren, C.M. 2014. Sexual development, size at maturity, size at maternity and fecundity of the blue shark \emph{{Prionace} glauca} ({Linnaeus}, 1758) in the {Southwest} {Atlantic}. Fish. Res. \textbf{160}: 18--32. doi: \url{https://doi.org/10.1016/j.fishres.2014.03.003}.

@@

\bibitem[\citeproctext]{ref-moulton_age_1992}
Moulton, P.L., Walker, T.I., and Saddlier, S.R. 1992. Age and growth-studies of gummy shark, \emph{{Mustelus} antarcticus} {Gunther}, and school shark, \emph{{Galeorhinus} galeus}({Linnaeus}), from {Southern} {Australian} waters. Aust. J. Mar. Fresh. Res. \textbf{43}: 1241--1267. doi: \href{https://doi.org/10.1071/MF9921241}{10.1071/MF9921241}.

\bibitem[\citeproctext]{ref-natanson_presence_2019}
Natanson, L.J., Deacy, B.M., Joyce, W., and Sulikowski, J. 2019. Presence of a resting population of female porbeagles ({Lamna} nasus), indicating a biennial reproductive cycle, in the western {North} {Atlantic} {Ocean}. Fish. B-noaa. \textbf{117}.

@@

\bibitem[\citeproctext]{ref-punt_estimating_1997}
Punt, A.E., Kennedy, R.B., and Frusher, S.D. 1997. Estimating the size-transition matrix for {Tasmanian} rock lobster, \emph{{Jasus} edwardsii}. Mar. Freshwater Res. \textbf{48}(8): 981--992. Available from \url{https://doi.org/10.1071/MF97017}.

@@

\bibitem[\citeproctext]{ref-quinn_quantitative_1999}
Quinn, I.I., and Deriso, R.B. 1999. Quantitative fish dynamics. Oxford University Press, New York, NY.

@@

\bibitem[\citeproctext]{ref-rambahiniarison_life_2018}
Rambahiniarison, J.M., Lamoste, M.J., Rohner, C.A., Murray, R., Snow, S., Labaja, J., Araujo, G., and Ponzo, A. 2018. Life history, growth, and reproductive biology of four mobulid species in the {Bohol} {Sea}, {Philippines}. Frontiers in Marine Science \textbf{5}: 269. doi: \href{https://doi.org/10.3389/fmars.2018.00269}{10.3389/fmars.2018.00269}.

\bibitem[\citeproctext]{ref-rigby_patterns_2015}
Rigby, C., and Simpfendorfer, C.A. 2015. Patterns in life history traits of deep-water chondrichthyans. Deep Sea Research Part II: Topical Studies in Oceanography \textbf{115}: 30--40. doi: \url{https://doi.org/10.1016/j.dsr2.2013.09.004}.

\bibitem[\citeproctext]{ref-roa_estimation_1999}
Roa, R., Ernst, B., and Tapia, F. 1999. Estimation of size at sexual maturity: An evaluation of analytical and resampling procedures. Fish. B-noaa. \textbf{97}: 570--580.

@@

\bibitem[\citeproctext]{ref-sato_how_2016}
Sato, K., Nakamura, M., Tomita, T., Toda, M., Miyamoto, K., and Nozu, R. 2016. How great white sharks nourish their embryos to a large size: Evidence of lipid histotrophy in lamnoid shark reproduction. Biology Open \textbf{5}(9): 1211--1215. doi: \href{https://doi.org/10.1242/bio.017939}{10.1242/bio.017939}.

\bibitem[\citeproctext]{ref-sedar_sedar_2012}
SEDAR. 2012. {SEDAR} 29 stock assessment report: {HMS} {Gulf} of {Mexico} blacktip shark. North Charleston, SC.

\bibitem[\citeproctext]{ref-sedar_sedar_2017}
SEDAR. 2017. {SEDAR} 54 stock assessment report: {HMS} sandbar shark. North Charleston, SC.

@@

\bibitem[\citeproctext]{ref-10.1371ux2fjournal.pone.0246734}
Smart, J.J., and Grammer, G.L. 2021. Modernising fish and shark growth curves with bayesian length-at-age models. Plos One \textbf{16}(2): 1--21. doi: \href{https://doi.org/10.1371/journal.pone.0246734}{10.1371/journal.pone.0246734}.

@@

\bibitem[\citeproctext]{ref-swenson2024}
Swenson, J.D., Brooks, E.N., Kacev, D., Boyd, C., Kinney, M.J., Marcy-Quay, B., Sévêque, A., Feldheim, K.A., and Komoroske, L.M. 2024. Accounting for unobserved population dynamics and aging error in close{-}kin mark{-}recapture assessments. Ecology and Evolution \textbf{14}(2). doi: \href{https://doi.org/10.1002/ece3.10854}{10.1002/ece3.10854}.

\bibitem[\citeproctext]{ref-taylor_stockrecruitment_2013}
Taylor, I.G., Gertseva, V., Methot, R.D., and Maunder, M.N. 2013. A stock--recruitment relationship based on pre-recruit survival, illustrated with application to spiny dogfish shark. Fish. Res. \textbf{142}: 15--21. doi: \url{https://doi.org/10.1016/j.fishres.2012.04.018}.

@@

\bibitem[\citeproctext]{ref-walker_fishery_1992}
Walker, T. 1992. Fishery simulation model for sharks applied to the {Gummy} {Shark}, \emph{{Mustelus} antarcticus} {Gunther}, from {Southern} {Australian} waters. Mar. Freshwater Res. \textbf{43}(1): 195--212. Available from \url{https://doi.org/10.1071/MF9920195}.

@@

\bibitem[\citeproctext]{ref-zuur_analyzing_2007}
Zuur, A., Ieno, E.N., and Smith, G.M. 2007. Analyzing ecological data. New York, NY.

>>>>

\[
C_{l} = \sum_a(F_{l}/Z_{l})(1-\exp(-Z_{l}))N_{l,a}
\]

@@

When using the 3PLF methods model performance typically declined as the periodicity of the reproductive cycle increased. This is seen in the positive correlation between \(\widehat{L'_{50}}\) and \(\widehat{P_{Max}}\), whereby true underlying \(P_{Max}\) affected the magnitude and direction of bias in these parameters (Figure 2). With lower values of \(P_{Max}\) it appeared to be more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote. As true underlying \(P_{Max}\) decreased there was also a greater potential for positive bias in \(\widehat{P_{Max}}\). Additionally, longer reproductive cycles resulted in fewer females in maternal condition, which was associated with larger bias (Figure 2).

@@

While the 3PLF-fixed method was in some cases able to reduce bias at lower sample sizes, a trade off in the use of this method was poorer interval coverage (Figure 4, Figure S6). For most simulations, interval coverage for \(\widehat{L'_{50}}\) was well below the expected level of 50\%, irrespective of sample size. In contrast, for the 3PLF-estimated method interval coverage oscillated around 50\% for both \(\widehat{L'_{50}}\) and \(\widehat{P_{Max}}\).

@@

Maternal data were re-analysed for 1087 sandbar sharks including 640 mature individuals of which 32\% were in maternal condition. Using the 3PLF-estimated method, the maximum likelihood estimate for \(\widehat{P_{Max}}\) was 0.48 (Table 2). Despite having a sample size of \textgreater{} 600 mature females, the proportion of maternal individuals at length was still uncertain and \(\widehat{P_{Max}}\) was estimated to lie between 0.39 and 0.60 with 95\% confidence (Figure 7). For 3PLF-fixed models a value of \(P_{Max} = 0.5\) was strongly supported (\(\Delta_{AIC}\) = 14.15) over the alternative fixed value of \(P_{Max} = 0.33\). The model with \(P_{Max} = 0.5\) also outperformed the 3PLF-estimated method (\(\Delta_{AIC}\) = 1.86). Given the fixed model had one fewer estimated parameter, both models had essentially the same level of support given the data (Burnham and Anderson 2002).

@@

The empirical analysis carried out using data for the sandbar shark illustrates how novel insights can be gained from maternal data by estimating \(P_{Max}\). Detailed reproductive studies of the western North Atlantic population by Baremore and Hale (2012) and Piercy et al. (2016) were ambiguous about the frequency of reproduction. In both studies, the low proportion of pregnant females and bimodality in ovarian follicle size led the authors to conclude that the reproductive cycle was most commonly triennial or possibly longer at the population level. Reanalysis of these combined data sets using the 3PLF-estimated method, however, resulted in a maximum likelihood estimate of \(\widehat{P_{Max}} = 0.48\), considerably larger than 0.37 - the proportion of pregnant mature females observed by Baremore and Hale (2012). Similarly, based on AIC values, a fixed value of \(P_{Max} = 0.5\) had a much higher relative plausibility than a value of 0.33 (Table 2).

This unexpected result may be related to the gradual attainment of asymptotic maternal status. Maturity in sandbar sharks appears to occur over an extended size range. \(L_{5}\) and \(L_{95}\) are 146cm and 176cm -- a length interval that corresponds to around seven years given the sandbar shark's slow growth rate (Hale and Baremore 2013). With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently, while larger sharks may ultimately be capable of reproducing biennially. Given that few females appear to attain this size (Hale and Baremore 2013), the modal reproductive frequency is still likely at least triennial within the population. Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management; not only are larger females more fecund (Baremore and Hale 2012), they also reproduce more frequently. This implies that size selective fishing or the removal of larger and older female sharks could have a disproportionately greater impact on population productivity.

@@

\section{Funding statement}\label{funding-statement}

The authors declare no specific funding for this work.

@@

\bibitem[\citeproctext]{ref-higgs_reproductive_2020}
Higgs, J.M., Hoffmayer, E.R., Sulikowski, J.A., Driggers, W.B., Stiller, D.A., and Hendon, J.M. 2020. Reproductive biology of the finetooth shark (\emph{{Carcharhinus} isodon}) in the northern {Gulf} of {Mexico}, with evidence of both annual and biennial reproduction. Mar. Freshwater Res. doi: \href{https://doi.org/10.1071/MF20120}{10.1071/MF20120}.

@@

\bibitem[\citeproctext]{ref-huveneers_future_2018}
Huveneers, C., Apps, K., Becerril-García, E.E., Bruce, B., Butcher, P.A., Carlisle, A.B., Chapple, T.K., Christiansen, H.M., Cliff, G., Curtis, T.H., Daly-Engel, T.S., Dewar, H., Dicken, M.L., Domeier, M.L., Duffy, C.A.J., Ford, R., Francis, M.P., French, G.C.A., Galván-Magaña, F., García-Rodríguez, E., Gennari, E., Graham, B., Hayden, B., Hoyos-Padilla, E.M., Hussey, N.E., Jewell, O.J.D., Jorgensen, S.J., Kock, A.A., Lowe, C.G., Lyons, K., Meyer, L., Oelofse, G., Oñate-González, E.C., Oosthuizen, H., O'Sullivan, J.B., Ramm, K., Skomal, G., Sloan, S., Smale, M.J., Sosa-Nishizaki, O., Sperone, E., Tamburin, E., Towner, A.V., Wcisel, M.A., Weng, K.C., and Werry, J.M. 2018. Future research directions on the {``{Elusive}''} white shark. Front. Mar. Sci. \textbf{5}: 455. doi: \href{https://doi.org/10.3389/fmars.2018.00455}{10.3389/fmars.2018.00455}.

@@

\bibitem[\citeproctext]{ref-montealegre-quijano_sexual_2014}
Montealegre-Quijano, S., Cardoso, A.T.C., Silva, R.Z., Kinas, P.G., and Vooren, C.M. 2014. Sexual development, size at maturity, size at maternity and fecundity of the blue shark \emph{{Prionace} glauca} ({Linnaeus}, 1758) in the {Southwest} {Atlantic}. Fish. Res. \textbf{160}: 18--32. doi: \href{https://doi.org/10.1016/j.fishres.2014.03.003}{10.1016/j.fishres.2014.03.003}.

@@

\bibitem[\citeproctext]{ref-moulton_age_1992}
Moulton, P.L., Walker, T.I., and Saddlier, S.R. 1992. Age and growth-studies of gummy shark, \emph{{Mustelus} antarcticus} {Gunther}, and school shark, \emph{{Galeorhinus} galeus} ({Linnaeus}), from {Southern} {Australian} waters. Aust. J. Mar. Fresh. Res. \textbf{43}: 1241--1267. doi: \href{https://doi.org/10.1071/MF9921241}{10.1071/MF9921241}.

\bibitem[\citeproctext]{ref-natanson_presence_2019}
Natanson, L.J., Deacy, B.M., Joyce, W., and Sulikowski, J. 2019. Presence of a resting population of female porbeagles (\emph{{Lamna} nasus}), indicating a biennial reproductive cycle, in the western {North} {Atlantic} {Ocean}. Fish. Bull. \textbf{117}.

@@

\bibitem[\citeproctext]{ref-punt_estimating_1997}
Punt, A.E., Kennedy, R.B., and Frusher, S.D. 1997. Estimating the size-transition matrix for {Tasmanian} rock lobster, \emph{{Jasus} edwardsii}. Mar. Freshwater Res. \textbf{48}(8): 981--992. doi: \href{https://doi.org/10.1071/MF97017}{10.1071/MF97017}.

@@

\bibitem[\citeproctext]{ref-quinn_quantitative_1999}
Quinn, I.I., and Deriso, R.B. 1999. Quantitative {Fish} {Dynamics}. Oxford University Press, New York, NY.

@@

\bibitem[\citeproctext]{ref-rambahiniarison_life_2018}
Rambahiniarison, J.M., Lamoste, M.J., Rohner, C.A., Murray, R., Snow, S., Labaja, J., Araujo, G., and Ponzo, A. 2018. Life history, growth, and reproductive biology of four mobulid species in the {Bohol} {Sea}, {Philippines}. Front. Mar. Sci. \textbf{5}: 269. doi: \href{https://doi.org/10.3389/fmars.2018.00269}{10.3389/fmars.2018.00269}.

\bibitem[\citeproctext]{ref-rigby_patterns_2015}
Rigby, C., and Simpfendorfer, C.A. 2015. Patterns in life history traits of deep-water chondrichthyans. Deep Sea Research Part II: Topical Studies in Oceanography \textbf{115}: 30--40. doi: \href{https://doi.org/10.1016/j.dsr2.2013.09.004}{10.1016/j.dsr2.2013.09.004}.

\bibitem[\citeproctext]{ref-roa_estimation_1999}
Roa, R., Ernst, B., and Tapia, F. 1999. Estimation of size at sexual maturity: An evaluation of analytical and resampling procedures. Fish. Bull. \textbf{97}: 570--580.

@@

\bibitem[\citeproctext]{ref-sato_how_2016}
Sato, K., Nakamura, M., Tomita, T., Toda, M., Miyamoto, K., and Nozu, R. 2016. How great white sharks nourish their embryos to a large size: Evidence of lipid histotrophy in lamnoid shark reproduction. Biol. Open \textbf{5}(9): 1211--1215. doi: \href{https://doi.org/10.1242/bio.017939}{10.1242/bio.017939}.

\bibitem[\citeproctext]{ref-sedar_sedar_2012}
SEDAR. 2012. {SEDAR} 29 {Stock} {Assessment} {Report}: {HMS} {Gulf} of {Mexico} {Blacktip} {Shark}. North Charleston, SC.

\bibitem[\citeproctext]{ref-sedar_sedar_2017}
SEDAR. 2017. {SEDAR} 54 {Stock} {Assessment} {Report}: {HMS} {Sandbar} {Shark}. North Charleston, SC.

@@

\bibitem[\citeproctext]{ref-10.1371ux2fjournal.pone.0246734}
Smart, J.J., and Grammer, G.L. 2021. Modernising fish and shark growth curves with {Bayesian} length-at-age models. Plos One \textbf{16}(2): 1--21. doi: \href{https://doi.org/10.1371/journal.pone.0246734}{10.1371/journal.pone.0246734}.

@@

\bibitem[\citeproctext]{ref-swenson2024}
Swenson, J.D., Brooks, E.N., Kacev, D., Boyd, C., Kinney, M.J., Marcy-Quay, B., Sévêque, A., Feldheim, K.A., and Komoroske, L.M. 2024. Accounting for unobserved population dynamics and aging error in close{-}kin mark{-}recapture assessments. Ecol. Evol. \textbf{14}(2). doi: \href{https://doi.org/10.1002/ece3.10854}{10.1002/ece3.10854}.

\bibitem[\citeproctext]{ref-taylor_stockrecruitment_2013}
Taylor, I.G., Gertseva, V., Methot, R.D., and Maunder, M.N. 2013. A stock--recruitment relationship based on pre-recruit survival, illustrated with application to spiny dogfish shark. Fish. Res. \textbf{142}: 15--21. doi: \href{https://doi.org/10.1016/j.fishres.2012.04.018}{10.1016/j.fishres.2012.04.018}.

@@

\bibitem[\citeproctext]{ref-walker_fishery_1992}
Walker, T. 1992. Fishery simulation model for sharks applied to the {Gummy} {Shark}, \emph{{Mustelus} antarcticus} {Gunther}, from {Southern} {Australian} waters. Mar. Freshwater Res. \textbf{43}(1): 195--212. doi: \href{https://doi.org/10.1071/MF9920195}{10.1071/MF9920195}.

@@

\bibitem[\citeproctext]{ref-zuur_analyzing_2007}
Zuur, A., Ieno, E.N., and Smith, G.M. 2007. Analyzing {Ecological} {Data}. New York, NY.
